---
title: "Assigment-2: R and SQL"
output: html_document
date: "2024-09-12"
---

#### Introduction
To gather insights on recent popular movies, I will choose six recently released films that have gained significant attention. Then, I will ask at least five people I know—friends, family, classmates, or even imaginary friends if needed—to rate each movie they have watched on a scale from 1 to 5.

```{r reading-from-mysql, include=TRUE}

install.packages("RMySQL", repos = "https://cran.r-project.org")
library(RMySQL)

user <- Sys.getenv("MYSQL_USER")
password <- Sys.getenv("MYSQL_PASSWORD")
host <- Sys.getenv("MYSQL_HOST")
dbname <- Sys.getenv("MYSQL_DBNAME")

conn <- dbConnect(MySQL(), user = user, password = password, host = host, dbname = dbname)

query <- "select * from movie_rankings"
movies_ranking_df <- dbGetQuery(conn, query)

query <- "select * from movies"
movies_df <- dbGetQuery(conn, query)

dbDisconnect(conn)
```

The first six rows of my friends’ ranking table records.

```{r displaying-movie-ranking-records}
library(knitr)
kable(head(movies_ranking_df))
```

The first six rows of my movies table records.

```{r displaying-all-movies}
kable(head(movies_df))
```

The average rating of my frinds for the following 6 movies is: 

```{r friends-ranking-displayed}
library(dplyr)

mean_ratings_df <- movies_ranking_df %>% group_by(movie_name) %>% 
  summarise(mean_rating = mean(movie_ranking, na.rm = TRUE))

library(ggplot2)
ggplot(mean_ratings_df, aes(x = movie_name, y = mean_rating, fill = movie_name)) + 
  geom_bar(stat = "identity") + 
  labs(x = "Movie Name", y = "Friends' Average Rating") +
  theme_minimal() 
```

#### Note: Missing Data in the Data Frame Example
Code example: "na.rm = true"is an argument used in functions to specify that missing values should be removed before performing the operation. This is particularly useful in functions like mean(), sum(), sd(), etc. 

Example vector with missing values:
data <- myNumbers(1, 2, NA, 4, 5)

Calculate the mean, removing NA values:
mean_value <- mean(data, na.rm = TRUE)

Its output will be 3

***I believe there are some other ways to do it, but this is the only one I was able to actually use in this use case.  

#### Back to our movies...
Normalized ratings representation up to 5 for the same movies according to IMDb.

```{r imdb-ranking-displayed}
normalized_imbd_rating_df = movies_df %>% group_by(movie_name) %>%
  summarise(mean_rating = mean(round((imdb_ranking / 10) * 5), na.rm = TRUE))

ggplot(normalized_imbd_rating_df, aes(x = movie_name, y = mean_rating, fill = movie_name)) + 
  geom_bar(stat = "identity") + 
  labs(x = "Movie Name", y = "imdb Average Rating") +
  theme_minimal() + 
  theme(axis.text.x = element_blank()) #make it look clean
```

There is a representation of now my frends' ranking is different from imdb ranking: 

```{r imdb-vs-friends-ranking-displayed}
library(tidyr)

#only common values
joined_movies <- inner_join(mean_ratings_df, normalized_imbd_rating_df, 
                            by = "movie_name", 
                            suffix = c(".mean_ratings_df", ".normalized_imbd_rating_df")) # which column from which table


# long format
long_movies <- joined_movies %>%
  pivot_longer(cols = starts_with("mean_rating"), 
               names_to = "source", 
               values_to = "rating") %>%
  mutate(source = recode(source, 
                         "mean_rating.mean_ratings_df" = "Friends", 
                         "mean_rating.normalized_imbd_rating_df" = "IMDb"))

#data plot
ggplot(long_movies, aes(x = movie_name, y = rating, fill = source)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_minimal() +
  labs(title = "Comparison of Friends' vs IMDb Rankings",
       x = "Movie Name",
       y = "Average Rating") +
  scale_fill_manual(name = "Rating Source", values = c('Friends' = '#E75480', 'IMDb' = 'blue'))

```

#### Colclusion:
Based on two graph graph “Friends’ Ranking vs IMDb Rankings,” we can conclude that:

 - The movies “Back to Back” and “Joker: Folie à Deux” received higher ratings from the friends’ group compared to IMDb.
 - “It Ends With Us” was rated higher by IMDb than by the friends’ group.
 - “Nowhere” and “The Watchers” have similar ratings from both sources.
This comparison shows how personal opinions from a group of friends can differ from the broader public consensus on IMDb. It highlights the subjective nature of movie ratings and how they can vary depending on the audience.

