---
title: "Assignment 7"
output: html_document
date: "2024-10-15"
author: "Inna Yedzinovich"
---

### Introduction
##### Assignment: working with JSON, HTML, XML, and Parquet in R 

We have received the following data from CUNYMart, located at 123 Example Street,
Anytown, USA.

Ask: "This data will be used for inventory analysis at the retailer. You are required to prepare the data
for analysis by formatting it in JSON, HTML, XML, and Parquet. Additionally, provide the pros
and cons of each format."

Let’s create a dataframe from the provided data:

```{r}
library(tibble)

electronics_df <- tibble::tribble(
  ~Category, ~Item_Name, ~Item_ID, ~Brand, ~Price, ~Series_ID, ~Series_Details,
  "Electronics", "Smartphone", 101, "TechBrand", 699.99, "101-A", "Color: Black, Storage: 64GB",
  "Electronics", "Smartphone", 101, "TechBrand", 699.99, "101-B", "Color: White, Storage: 128GB",
  "Electronics", "Laptop", 102, "CompuBrand", 1099.99, "102-A", "Color: Silver, Storage: 256GB",
  "Electronics", "Laptop", 102, "CompuBrand", 1099.99, "102-B", "Color: Space Gray, Storage: 512GB",
  "Home Appliances", "Refrigerator", 201, "HomeCool", 899.99, "201-A", "Color: Stainless Steel, Capacity: 20 cu ft",
  "Home Appliances", "Refrigerator", 201, "HomeCool", 899.99, "201-B", "Color: White, Capacity: 18 cu ft",
  "Home Appliances", "Washing Machine", 202, "CleanTech", 499.99, "202-A", "Type: Front Load, Capacity: 4.5 cu ft",
  "Home Appliances", "Washing Machine", 202, "CleanTech", 499.99, "202-B", "Type: Top Load, Capacity: 5.0 cu ft",
  "Clothing", "T-Shirt", 301, "FashionCo", 19.99, "301-A", "Color: Blue, Size: S",
  "Clothing", "T-Shirt", 301, "FashionCo", 19.99, "301-B", "Color: Red, Size: M",
  "Clothing", "T-Shirt", 301, "FashionCo", 19.99, "301-C", "Color: Green, Size: L",
  "Clothing", "Jeans", 302, "DenimWorks", 49.99, "302-A", "Color: Dark Blue, Size: 32",
  "Clothing", "Jeans", 302, "DenimWorks", 49.99, "302-B", "Color: Light Blue, Size: 34",
  "Books", "Fiction Novel", 401, "-", 14.99, "401-A", "Format: Hardcover, Language: English",
  "Books", "Fiction Novel", 401, "-", 14.99, "401-B", "Format: Paperback, Language: Spanish",
  "Books", "Non-Fiction Guide", 402, "-", 24.99, "402-A", "Format: eBook, Language: English",
  "Books", "Non-Fiction Guide", 402, "-", 24.99, "402-B", "Format: Paperback, Language: French",
  "Sports Equipment", "Basketball", 501, "SportsGear", 29.99, "501-A", "Size: Size 7, Color: Orange",
  "Sports Equipment", "Tennis Racket", 502, "RacketPro", 89.99, "502-A", "Material: Graphite, Color: Black",
  "Sports Equipment", "Tennis Racket", 502, "RacketPro", 89.99, "502-B", "Material: Aluminum, Color: Silver"
)

print(electronics_df)
```

Generate a JSON file from this dataset.

```{r}
library(jsonlite)
json_data <- toJSON(electronics_df, pretty = TRUE)
write(json_data, file = "electronics.json")
```

Save the JSON file in github. 
GitHub Link: https://raw.githubusercontent.com/Yedzinovich/Data-607/refs/heads/main/electronics.json

####1 - JSON
Let’s load our generated JSON file into a dataframe named electronics_json.

```{r}
library(jsonlite)
electronics_json <- fromJSON("https://raw.githubusercontent.com/Yedzinovich/Data-607/refs/heads/main/electronics.json")
print(electronics_json)
```

####2 - HTML
I will use a JSON file and a format converter to transform it into HTML format. Let’s then read our generated HTML file into a dataframe named electronics_html. 

```{r}
library(rvest)
library(dplyr)

url <- "https://raw.githubusercontent.com/Yedzinovich/Data-607/refs/heads/main/electronics_html.html"
html_content <- read_html(url)

electronics_html <- html_content %>%
  html_node("table") %>%
  html_table(fill = TRUE)
print(electronics_html)
```

#### 3- XML
I will use a JSON file and a format converter to transform it into XML format. Let’s then read our generated XML file into a dataframe named electronics_xml. 

```{r}
library(xml2)
url <- "https://raw.githubusercontent.com/Yedzinovich/Data-607/refs/heads/main/electronics_xml.xml"
xml_content <- read_xml(url)
items <- xml_find_all(xml_content, "//item")
print(items)

categories <- xml_text(xml_find_all(items, "Category"))
item_names <- xml_text(xml_find_all(items, "Item_Name"))
item_ids <- xml_text(xml_find_all(items, "Item_ID"))
brands <- xml_text(xml_find_all(items, "Brand"))
prices <- xml_text(xml_find_all(items, "Price"))
series_ids <- xml_text(xml_find_all(items, "Series_ID"))
series_details <- xml_text(xml_find_all(items, "Series_Details"))

electronics_xml <- data.frame(
  Category = categories,
  Item_Name = item_names,
  Item_ID = item_ids,
  Brand = brands,
  Price = prices,
  Series_ID = series_ids,
  Series_Details = series_details,
  stringsAsFactors = FALSE
)

print(electronics_xml)
```

#### 4 - Parquet
I will use a JSON file and a format converter to transform it into Parquet format. Let’s then read our generated Parquet file into a dataframe named electronics_parquet. 

```{r}
library(arrow)
url <- "https://raw.githubusercontent.com/Yedzinovich/Data-607/refs/heads/main/electronics.parquet"

temp_file <- tempfile(fileext = ".parquet")
download.file(url, temp_file, mode = "wb")

electronics_parquet <- read_parquet(temp_file)
print(electronics_parquet)
```

### Conclusion 

#### JSON (JavaScript Object Notation)
  Pros: Easy to read, lightweight, widely supported.
  Cons: Can be large, no schema, limited data types.
  
#### HTML (HyperText Markup Language)
Pros: Great for web pages, rich formatting.
Cons: Not for data storage, bulky, limited structure.

#### XML (eXtensible Markup Language)
Pros: Self-descriptive, flexible, supports schema.
Cons: Very large, complex to parse, slower.

#### Parquet
Pros: Efficient storage, good compression, fast for big data.
Cons: Not human-readable, complex tools needed, less support.

In conclusion, choosing the right data format depends on our specific needs and context. JSON is ideal for lightweight data interchange due to its simplicity and readability. HTML excels in presenting data on web pages with rich formatting options. XML offers flexibility and self-descriptive tags, making it suitable for complex hierarchical data structures, though it can be verbose and slower to process. Parquet stands out for its efficiency in storing and processing large-scale data, especially in analytical scenarios, despite requiring more sophisticated tools and being less human-readable.

Each format has its unique strengths and weaknesses, making them suitable for different applications. 
