---
title: "Assignment 7"
output: html_document
date: "2024-10-15"
author: "Inna Yedzinovich"
---

### Introduction
##### Assignment: working with JSON, HTML, XML, and Parquet in R 

We have received the following data from CUNYMart, located at 123 Example Street,
Anytown, USA.

Ask: "This data will be used for inventory analysis at the retailer. You are required to prepare the data
for analysis by formatting it in JSON, HTML, XML, and Parquet. Additionally, provide the pros
and cons of each format."

Let’s create a dataframe from the provided data:

```{r}
library(tibble)

electronics_df <- tibble::tribble(
  ~Category, ~Item_Name, ~Item_ID, ~Brand, ~Price, ~Series_ID, ~Series_Details,
  "Electronics", "Smartphone", 101, "TechBrand", 699.99, "101-A", "Color: Black, Storage: 64GB",
  "Electronics", "Smartphone", 101, "TechBrand", 699.99, "101-B", "Color: White, Storage: 128GB",
  "Electronics", "Laptop", 102, "CompuBrand", 1099.99, "102-A", "Color: Silver, Storage: 256GB",
  "Electronics", "Laptop", 102, "CompuBrand", 1099.99, "102-B", "Color: Space Gray, Storage: 512GB",
  "Home Appliances", "Refrigerator", 201, "HomeCool", 899.99, "201-A", "Color: Stainless Steel, Capacity: 20 cu ft",
  "Home Appliances", "Refrigerator", 201, "HomeCool", 899.99, "201-B", "Color: White, Capacity: 18 cu ft",
  "Home Appliances", "Washing Machine", 202, "CleanTech", 499.99, "202-A", "Type: Front Load, Capacity: 4.5 cu ft",
  "Home Appliances", "Washing Machine", 202, "CleanTech", 499.99, "202-B", "Type: Top Load, Capacity: 5.0 cu ft",
  "Clothing", "T-Shirt", 301, "FashionCo", 19.99, "301-A", "Color: Blue, Size: S",
  "Clothing", "T-Shirt", 301, "FashionCo", 19.99, "301-B", "Color: Red, Size: M",
  "Clothing", "T-Shirt", 301, "FashionCo", 19.99, "301-C", "Color: Green, Size: L",
  "Clothing", "Jeans", 302, "DenimWorks", 49.99, "302-A", "Color: Dark Blue, Size: 32",
  "Clothing", "Jeans", 302, "DenimWorks", 49.99, "302-B", "Color: Light Blue, Size: 34",
  "Books", "Fiction Novel", 401, "-", 14.99, "401-A", "Format: Hardcover, Language: English",
  "Books", "Fiction Novel", 401, "-", 14.99, "401-B", "Format: Paperback, Language: Spanish",
  "Books", "Non-Fiction Guide", 402, "-", 24.99, "402-A", "Format: eBook, Language: English",
  "Books", "Non-Fiction Guide", 402, "-", 24.99, "402-B", "Format: Paperback, Language: French",
  "Sports Equipment", "Basketball", 501, "SportsGear", 29.99, "501-A", "Size: Size 7, Color: Orange",
  "Sports Equipment", "Tennis Racket", 502, "RacketPro", 89.99, "502-A", "Material: Graphite, Color: Black",
  "Sports Equipment", "Tennis Racket", 502, "RacketPro", 89.99, "502-B", "Material: Aluminum, Color: Silver"
)

print(electronics_df)
```

Generate a JSON file from this dataset.

```{r}
library(jsonlite)
json_data <- toJSON(electronics_df, pretty = TRUE)
write(json_data, file = "electronics.json")
```

Save the JSON file in github. 
GitHub Link: https://raw.githubusercontent.com/Yedzinovich/Data-607/refs/heads/main/electronics.json

Let’s load our generated JSON file into a dataframe named electronics_json.

```{r}
library(jsonlite)
electronics_json <- fromJSON("https://raw.githubusercontent.com/Yedzinovich/Data-607/refs/heads/main/electronics.json")
print(electronics_json)
```

I am going to use JSON file and format converter to make it into HTML format. 
Let's read from out generated HTML file into a data frame - electronics_html


```{r}
library(jsonlite)
electronics_json <- fromJSON("https://raw.githubusercontent.com/Yedzinovich/Data-607/refs/heads/main/electronics.json")
print(electronics_json)
```
