---
title: "Data 607 Project 3"
output: html_document
author: "Md. Simon Chowdhury, Inna Yedzinovich, Zaneta Paulusova, Md Asaduzzaman, Md. Asadul"
date: "`r Sys.Date()`"

---
## Introduction 

For this project we focused on answering the question “What are the most valued data science skills?”.

Our approach to answering this question involved creating a survey to identify five valuable data science skills as perceived by survey respondents. We then compared these findings to the article titled "Data Science Skills Survey 2022 – By AIM and Great Learning." In this article, we focused on a table that lists common skills desired by recruiters, categorized by years of experience. We will adjust the data to ensure the skills align with one another. Additionally, we will calculate the mean for the various years of experience columns from the website's data. Finally, we will combine both datasets into one graph for a comparative analysis of our internal survey results alongside the findings from the website.


#### Tables creation for our data

```{r}
library(DBI)
library(RMySQL)

user <- Sys.getenv("MYSQL_USER")
password <- Sys.getenv("MYSQL_PASSWORD")
host <- Sys.getenv("MYSQL_HOST")
dbname <- Sys.getenv("MYSQL_DBNAME")

conn <- dbConnect(RMySQL::MySQL(), user = user, password = password, host = host, dbname = dbname)

create_experience_table <- "
CREATE TABLE experience (
  experience_id INT NOT NULL,
  respondent_id INT NULL,
  data_science_experience TINYINT(1) NULL,
  software_engineering_experience TINYINT(1) NULL,
  PRIMARY KEY (experience_id)
);"

create_respondents_table <- "
CREATE TABLE respondents (
  respondent_id INT NOT NULL,
  first_name VARCHAR(45) NULL,
  last_name VARCHAR(45) NULL,
  age INT NULL,
  PRIMARY KEY (respondent_id)
);"

create_interestareas_table <- "
CREATE TABLE interestareas (
  interest_id INT NOT NULL,
  respondent_id INT NULL,
  interest_area VARCHAR(45) NULL,
  PRIMARY KEY (interest_id)
);"

create_softskills_table <- "
CREATE TABLE softskills (
  soft_skill_id INT NOT NULL,
  respondent_id INT NULL,
  soft_skill VARCHAR(45) NULL,
  PRIMARY KEY (soft_skill_id)
);"

create_programminglanguages_table <- "
CREATE TABLE programminglanguages (
  language_id INT NOT NULL,
  respondent_id INT NULL,
  language VARCHAR(45) NULL,
  PRIMARY KEY (language_id)
);"

create_learningresources_table <- "
CREATE TABLE learningresources (
  resource_id INT NOT NULL,
  respondent_id INT NULL,
  resource VARCHAR(45) NULL,
  PRIMARY KEY (resource_id)
);"

create_valuableskills_table <- "
CREATE TABLE valuableskills (
  valuable_skill_id INT NOT NULL,
  respondent_id INT NULL,
  skill_rank INT NULL,
  skill_name VARCHAR(45) NULL,
  PRIMARY KEY (valuable_skill_id)
);"

dbExecute(conn, create_experience_table)
dbExecute(conn, create_respondents_table)
dbExecute(conn, create_interestareas_table)
dbExecute(conn, create_softskills_table)
dbExecute(conn, create_programminglanguages_table)
dbExecute(conn, create_learningresources_table)
dbExecute(conn, create_valuableskills_table)

dbDisconnect(conn)

```

WRITE BETTER TODOOOO

Load the CSV into R
Tidy and Normalize the Data
Load the Data into the Database

```{r}
library(dplyr)
library(tidyr)

data <- read.csv("https://raw.githubusercontent.com/simonchy/DATA607/refs/heads/main/week%208/Cleaned_Augmented_Data_Science_Skills.csv")
colnames(data)
colnames(data) <- tolower(colnames(data))

data <- data %>%
  rename(
    timestamp = timestamp,
    first_name = first.name.or.nickname,
    valuable_skills = list.the.5.most.valuable.data.science.skills..separated.by.commas.,
    email = email.address,
    age = age,
    data_science_experience = any.data.science.data.analytics.experience.,
    software_engineering_experience = any.software.engineering.experience.,
    programming_languages = which.programming.languages.do.you.use.most.frequently.,
    learning_resources = what.resources.do.you.use.for.learning.new.data.science.skills.,
    interest_areas = what.areas.of.data.science.are.you.most.interested.in.learning.more.about.,
    skill_1 = name..1.most.most.valuable.data.science.skill,
    skill_2 = name..2.most.most.valuable.data.science.skill,
    skill_3 = name..3.most.most.valuable.data.science.skill,
    skill_4 = name..4.most.most.valuable.data.science.skill,
    skill_5 = name..5.most.most.valuable.data.science.skill,
    soft_skill = which.soft.skill.do.you.think.is.most.important.for.a.data.scientist.
  ) %>%
  mutate(across(everything(), tolower))

# Create respondent_id before separating rows
data <- data %>%
  mutate(respondent_id = row_number())

# Normalize the data
respondents <- data %>%
  select(first_name, age, respondent_id)

experience <- data %>%
  select(data_science_experience, software_engineering_experience, respondent_id) %>%
  mutate(experience_id = row_number())

valuable_skills <- data %>%
  select(skill_1, skill_2, skill_3, skill_4, skill_5, respondent_id) %>%
  pivot_longer(cols = starts_with("skill_"), names_to = "skill_rank", values_to = "skill_name") %>%
  mutate(valuable_skill_id = row_number())

programming_languages <- data %>%
  select(respondent_id, programming_languages) %>%
  separate_rows(programming_languages, sep = ",") %>%
  mutate(language_id = row_number())

learning_resources <- data %>%
  select(respondent_id, learning_resources) %>%
  separate_rows(learning_resources, sep = ",") %>%
  mutate(resource_id = row_number())

interest_areas <- data %>%
  select(respondent_id, interest_areas) %>%
  separate_rows(interest_areas, sep = ",") %>%
  mutate(interest_id = row_number())

soft_skills <- data %>%
  select(soft_skill, respondent_id) %>%
  mutate(soft_skill_id = row_number())

conn <- dbConnect(RMySQL::MySQL(), user = Sys.getenv("MYSQL_USER"), password = Sys.getenv("MYSQL_PASSWORD"), host = Sys.getenv("MYSQL_HOST"), dbname = Sys.getenv("MYSQL_DBNAME"))

# Load the data into the database
dbWriteTable(conn, "respondents", respondents, overwrite = TRUE, row.names = FALSE)
dbWriteTable(conn, "experience", experience, overwrite = TRUE, row.names = FALSE)
dbWriteTable(conn, "valuableskills", valuable_skills, overwrite = TRUE, row.names = FALSE)
dbWriteTable(conn, "programminglanguages", programming_languages, overwrite = TRUE, row.names = FALSE)
dbWriteTable(conn, "learningresources", learning_resources, overwrite = TRUE, row.names = FALSE)
dbWriteTable(conn, "interestareas", interest_areas, overwrite = TRUE, row.names = FALSE)
dbWriteTable(conn, "softskills", soft_skills, overwrite = TRUE, row.names = FALSE)

dbDisconnect(conn)

```

BETTER DESCRIPTION TODOOOOO


```{r}
library(ggplot2)

conn <- dbConnect(RMySQL::MySQL(), user = Sys.getenv("MYSQL_USER"), password = Sys.getenv("MYSQL_PASSWORD"), host = Sys.getenv("MYSQL_HOST"), dbname = Sys.getenv("MYSQL_DBNAME"))
valuable_skills <- dbReadTable(conn, "valuableskills")
dbDisconnect(conn)

skill_counts <- valuable_skills %>%
  count(skill_name, sort = TRUE)
print(skill_counts)

ggplot(skill_counts, aes(x = reorder(skill_name, n), y = n)) +
  geom_bar(stat = "identity", fill = "pink") +
  labs(title = "Most Valuable Data Science Skills", x = "Data Science Skills", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Let's make it look cleaner: 

```{r}

# Group similar skills together using regex grepl
valuable_skills <- valuable_skills %>%
  mutate(skill_group = case_when(
    grepl("machine learning|ML", skill_name, ignore.case = TRUE) ~ "Machine Learning",
    grepl("programming|coding|software|Python|R", skill_name, ignore.case = TRUE) ~ "Programming",
    grepl("statistics|statistical", skill_name, ignore.case = TRUE) ~ "Statistics",
    grepl("data|database", skill_name, ignore.case = TRUE) ~ "Data Management",
    grepl("math|algorithm", skill_name, ignore.case = TRUE) ~ "Mathematics & Algorithms",
    grepl("communication|presentation", skill_name, ignore.case = TRUE) ~ "Communication",
    grepl("teamwork|collaborat", skill_name, ignore.case = TRUE) ~ "Teamwork",
    grepl("critical thinking|problem solving|analysis", skill_name, ignore.case = TRUE) ~ "Critical Thinking",
    TRUE ~ "Other"
  ))

skill_counts <- valuable_skills %>%
  count(skill_group, sort = TRUE)

ggplot(skill_counts, aes(x = reorder(skill_group, n), y = n)) +
  geom_bar(stat = "identity", fill = "pink") +
  labs(title = "Most Valuable Data Science Skills", x = "Skill Categories", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


Now, we can compare our results with the findings from the Data Science Skills Survey 2022 by AIM and Great Learning, available at this link.

Below is a table from that website showing the common skills sought by recruiters across different years of experience. The aim is to determine if the skills identified in our internal survey align with those highlighted by recruiters on the website. We will be calculation the mean of the years of experience as we did not request this information on our survey.

```{r}
common_skills_years <- read.csv("https://raw.githubusercontent.com/ZanetaP02/DATA-607/refs/heads/main/Data%20Science%20Skills%20Survey%202022%20-%20By%20AIM%20and%20Great%20Learning%20-%20Common%20skills%20by%20years%20table.csv")
common_skills_years
```

## Rename columns fo essier use
```{r}
up_csy <-common_skills_years
colnames(common_skills_years)[1] <-"skill_group"
colnames(common_skills_years)[2] <-">3yrs_%"
colnames(common_skills_years)[3] <-"3-5yrs_%"
colnames(common_skills_years)[4] <-"6-10yrs_%"
colnames(common_skills_years)[5] <-"10yrs<_%"
update <-common_skills_years
colnames(common_skills_years) <- c("skill_group", "<3yrs_%", "3-5yrs_%", "6-10yrs_%", "10yrs<_%")
head(up_csy)
```

## moving the diffrent years of expericance columns to one column
```{r}
up_csy1 <- up_csy %>%
pivot_longer(cols = c('>3yrs_%','3-5yrs_%','6-10yrs_%','10yrs<_%'), names_to = "year_experiance", values_to = "percentage")
head(up_csy1)
```
 
## calculating the mean of the skills
```{r}
avg_csy <- up_csy1 %>% group_by(skill_group) %>% 
  summarise(mean_percentage=mean(percentage),
            .groups = 'drop')
avg_csy
```


## grouping skills to match internal survey
```{r}
avg_csy1 <- avg_csy %>%
 mutate(skill_group = case_when(
    grepl("communication skills", skill_group, ignore.case =TRUE) ~ "Communication",
    grepl("programming knowledge", skill_group, ignore.case =TRUE) ~ "Programming",
    grepl("statistics",  skill_group, ignore.case =TRUE) ~ "Statistics",
    grepl("machine learning",  skill_group, ignore.case =TRUE) ~ "Machine Learning",
    grepl("linear algebra & calculus",  skill_group, ignore.case =TRUE) ~ "Mathematics & Algorithms",
    grepl("big data",  skill_group, ignore.case =TRUE) ~ "Data Management",
    grepl("deep learning",  skill_group, ignore.case =TRUE) ~ "Critical Thinking",
    grepl("data visualisation|business acumen|presentation skills|model deployment|data wrangling and pre-processing|domain expertise",  skill_group, ignore.case =TRUE) ~ "Other",
  ))
avg_csy1
```
 
## calculating mean of grouping skills to match internal survey skills
```{r}
avg_skills <- avg_csy1 %>% group_by(skill_group) %>% 
  summarise(mean_percentage=mean(mean_percentage),
            .groups = 'drop')
avg_skills
```
 
 
```{r}
ggplot(avg_skills, aes(x = reorder(skill_group, mean_percentage), y = mean_percentage)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Data Science Skills", x = "Skill Categories", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Now, let’s align our survey skills data with the format used on the website by converting the survey results into percentages.

```{r}
total_count <- sum(skill_counts$n)
total_count

skill_counts <- skill_counts %>%
  mutate(percentage = (n / total_count) * 100)
skill_counts

ggplot(skill_counts, aes(x = reorder(skill_group, percentage), y = percentage)) +
  geom_bar(stat = "identity", fill = "pink") +
  labs(title = "Most Valuable Data Science Skills", x = "Skill Categories", y = "Percentage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r}
merged_skill_counts <- avg_skills %>%
  left_join(skill_counts, by = "skill_group")
merged_skill_counts

merged_skill_counts <- merged_skill_counts %>%
  select(-n) %>%
  rename(
    website = mean_percentage,
    survey = percentage
  )

merged_skill_counts

```
 

```{r}
in_ex_skills <- merged_skill_counts %>%
  pivot_longer(cols = c('website', 'survey'), names_to = "website/survey", values_to = "mean %")
in_ex_skills 
```

```{r}
ggplot(in_ex_skills, aes(x = skill_group, y = `mean %`, fill = `website/survey`)) +
  geom_col(position = position_dodge()) +
  ggtitle("Comparison of Skill Groups Between Internal Survey and External Data") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



### Conclusion

TODOOOOOOOOOOOO
